---
title: "R Practice"
author: "JAM"
date: "6/2/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setting wd}
library(rstudioapi)
setwd(dirname(getActiveDocumentContext()$path))
```

## Intro to markdown

Markdown allows you to do all things like **bold**, *italic*, `code font`, and more!

Make a list with asterisks like:

-   bold with double-asterisk
-   italics with underscores
-   code-type font with backticks (\`\`)

You can also use hyphens instead of asterisks.

For a numbered list:

1.  bold......
2.  italics...
3.  code......

Hyperlinks are like [Hyperlinked text](https://mylink.org)


Load gapminder data along with plotting package.

```{r load-data}
# This line will not run
library(dplyr)
library(gapminder) # This chunk of this line will not run
library(ggplot2)
gap <- gapminder
```


Plot the data

```{r make-plot}
plot(lifeExp ~ year, data = gap)
```


```{r}
temp <- c(1, 2, 3)
```


```{r make_cats}
cats <- data.frame(coat = c("calico", "black", "tabby"), 
                   weight= c(2.1, 5.0, 3.2), 
                   likes_string = c(1, 0, 1))
write.csv(x = cats, file="data/feline-data.csv", row.names=FALSE)
```

```{r read_cats}
cats2 <- read.csv("data/feline-data.csv", stringsAsFactors = TRUE)
str(cats)
cats$weight <- cats$weight + 1
```


Creating vector and coercing types

```{r}
my_vec <- vector(mode="character", length=3)
combine_vector <- c(1, 2, "bananas")
```

change names

```{r}
names(cats)[3] <- "likes_hunting"
```


Add columns and rows to a df
```{r}
cats <- read.csv("data/feline-data.csv")
age <- c(2, 3, 5)
cats <- cbind(cats, age)
newRow <- list("tortoiseshell", 3.3, TRUE, 9)
cats <- rbind(cats, newRow)
```


```{r}
human_age <- cats$age * 7
human_age <- as.factor(human_age)
human_age <- as.numeric(as.character(human_age))
human_age <- human_age / 7
```

```{r}
cats$coat <- as.factor(cats$coat)
cats <- rbind(cats, c("white", 3.4, 0, 1))
cats <- rbind(cats, c("white", 3.4, 0, 1))
cats <- cats[-(5:6),]

cats <- na.omit(cats)
# cats[,-4] Removes column

drop <- names(cats) %in% c("weight")
cats[,!drop]
```









