----
title: "R Practice"
author: "JAM"
date: "6/2/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setting wd}
library(rstudioapi)
setwd(dirname(getActiveDocumentContext()$path))
```

## Intro to markdown

Markdown allows you to do all things like **bold**, *italic*, `code font`, and more!

Make a list with asterisks like:

-   bold with double-asterisk
-   italics with underscores
-   code-type font with backticks (\`\`)

You can also use hyphens instead of asterisks.

For a numbered list:

1.  bold......
2.  italics...
3.  code......

Hyperlinks are like [Hyperlinked text](https://mylink.org)


Load gapminder data along with plotting package.

```{r load-data}
# This line will not run
library(dplyr)
library(gapminder) # This chunk of this line will not run
library(ggplot2)
gap <- gapminder
```


Plot the data

```{r make-plot}
plot(lifeExp ~ year, data = gap)
```


```{r}
temp <- c(1, 2, 3)
```


```{r make_cats}
cats <- data.frame(coat = c("calico", "black", "tabby"), 
                   weight= c(2.1, 5.0, 3.2), 
                   likes_string = c(1, 0, 1))
write.csv(x = cats, file="data/feline-data.csv", row.names=FALSE)
```

```{r read_cats}
cats2 <- read.csv("data/feline-data.csv", stringsAsFactors = TRUE)
str(cats)
cats$weight <- cats$weight + 1
```


Creating vector and coercing types

```{r}
my_vec <- vector(mode="character", length=3)
combine_vector <- c(1, 2, "bananas")
```

change names

```{r}
names(cats)[3] <- "likes_hunting"
```


Add columns and rows to a df
```{r}
cats <- read.csv("data/feline-data.csv")
age <- c(2, 3, 5)
cats <- cbind(cats, age)
newRow <- list("tortoiseshell", 3.3, TRUE, 9)
cats <- rbind(cats, newRow)
```


```{r}
human_age <- cats$age * 7
human_age <- as.factor(human_age)
human_age <- as.numeric(as.character(human_age))
human_age <- human_age / 7
```

```{r}
cats$coat <- as.factor(cats$coat)
cats <- rbind(cats, c("white", 3.4, 0, 1))
cats <- rbind(cats, c("white", 3.4, 0, 1))
cats <- cats[-(5:6),]

cats <- na.omit(cats)
# cats[,-4] Removes column

drop <- names(cats) %in% c("weight")
cats[,!drop]
```


```{r}
library(readxl)
gapminder <- read_excel("data/gapminder_data.xlsx")

```

Inspecting a dataframe

```{r}
summary(gapminder)
summary(gapminder$lifeExp)

gapminder$country <- factor(gapminder$country)
gapminder$continent <- factor(gapminder$continent)
summary(gapminder)

length(gapminder) #ncol(gapminder) works
nrow(gapminder)
colnames(gapminder)

head(gapminder, n = 6) #tail is opposite

gapminder[sample(nrow(gapminder), 10), ]

gapminder[1:10, 2:5]
```

```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c("a", "b", "c", "d", "e")

x[c(1, 3)] # discrete items
x[1:3] # slice
x[c(1, 1, 3)]
x[6] #NA - beyond the end
x[(-c(1, 4))]
x[-1:3] # this would go from -1, 0, 1, 2, 3 which is wrong
# by name
x[c("a", "c")]


x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c("a", "b", "c", "d", "e")
x[x>7 & x < 7.5] # and
x[x>7 | x < 7.5] # or

x2 <- 1:3
names(x2) <- c("a", "a" ,"a")
x2["a"] # duplicate names are possible but not advisible

x
x[!names(x) %in% c("a", "c")]


```

Select southeast Asian countries


```{r}
seAsia <- c("Myanmar", "Thialand", "Cambodia", "Vietnam", "Laos")
seAsia_gap <- gapminder[which(gapminder$country %in% seAsia), ]
```


```{r}
gap_60s <- gapminder[gapminder$year %in% 1960:1969,]
```


**PICTURES!!!!!**

```{r}
library(ggplot2)
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y=lifeExp)) + geom_point()
```

```{r}
# This creates the framework, but doesn't know what to plot
ggplot(data=gapminder, mapping = aes(x=gdpPercap, y = lifeExp))
```


```{r}
ggplot(data=gapminder, mapping = aes(x=year, y = lifeExp)) + geom_point()
```

Lines! 

```{r}
ggplot(data=gapminder, mapping = aes(x=year, y=lifeExp, color=continent)) + geom_line()
```

Lines that makes a little sense

```{r}
ggplot(data=gapminder, mapping=aes(x=year, y=lifeExp, by=country)) + geom_line(mapping = aes(color=continent)) + geom_point(color="yellow")
```


```{r}
logged_points <- ggplot(gapminder, mapping = aes(x=gdpPercap, y=lifeExp)) + geom_point() + scale_x_log10()
logged_points <- logged_points + geom_smooth(method='loess')
logged_points
```


```{r}
americas <- gapminder[gapminder$continent == "Americas",]

ggplot(data=americas, mapping = aes(x=year, y=lifeExp)) + geom_line() + facet_wrap(~ country) + theme(axis.text.x = element_text(angle=45))

ggsave(filename="results/lifeExp_americas.png")
```


Save Americas!

```{r}
write.csv(file = "clean_data/americas_gap.csv", x = americas, row.names=FALSE)
```

Dplyr!

```{r}
mean(gapminder[gapminder$continent == "Africa",]$gdpPercap)
```


```{r}
library(dplyr)
year_country_gdp <- select(gapminder, year, country, gdpPercap)
str(year_country_gdp)
smaller_gap <- gapminder %>% select(-continent) %>% rename(gdp_per_capita = gdpPercap)
```


```{r}
y_cntr_gdp_eur <- gapminder %>% filter(continent=="Europe") %>% select(year, country, gdpPercap)

```


Combine with grouping

```{r}
gapminder %>% group_by(continent, year) %>% summarize(mean_gdpPercap = mean(gdpPercap), 
                                                sd_gdpPercap = sd(gdpPercap), 
                                                mean_raw_gdp = mean(gdpPercap*pop), 
                                                tot_raw_gdp = sum(gdpPercap*pop))
```

Mutate!

```{r}
gapminder %>% mutate(gdp_billion = gdpPercap*pop/10^9) %>%
  group_by(continent, year) %>% 
  summarize(mean_gdp_billion = mean(gdp_billion))
```

Combine with ggplot
```{r}
gapminder %>% filter(continent =="Americas") %>% 
  ggplot(mapping = aes(x=year, y=lifeExp)) + geom_line() + facet_wrap( ~ country) + theme(axis.text.x = element_text(angle=45))
```







